---
title: "Copepod_Ms_Code"
output: html_document
editor_options: 
  chunk_output_type: console
---
Load in data and libraries
```{r}
CopeData <- read.csv(url("https://raw.githubusercontent.com/marisolgr/NEP_Copepods_ThermalPatterns/main/Ashlock_Copepod_data_29Jan2021.csv"))
CopeData

library(lme4)
library(lmtest)
library(MASS)
library(fmsb)
library(faraway)
library(rsq)
library(tidyverse)
```

Set Region as factor
```{r}
head(CopeData)
str(CopeData)
CopeData$region <- as.factor(CopeData$region)
str(CopeData)
```

Add presence absence columns
```{r}
CopeData$paCalanus <- ifelse(CopeData$Calanus.pacificus >0, 1, 0)

CopeData$paOithona <- ifelse(CopeData$Oithona.spp. >0, 1, 0)

CopeData$paNeocalanus <- ifelse(CopeData$Neocalanus.plumchrus.V >0, 1, 0)

head(CopeData)
```

```{r}
CopeData<- CopeData[complete.cases(CopeData), ]
str(CopeData$paNeocalanus)
str(CopeData$paOithona)
str(CopeData$paCalanus)
head(CopeData)
tail(CopeData)
```

Create MonthDay column
```{r}
CopeData$MonthDay <- CopeData$month+((CopeData$day)/30.5)
```

Rename date column
```{r}
colnames(CopeData)[colnames(CopeData)=="MonthDay"] <- "date"
```


Modify temperature to Celcius from Kelvin
```{r}
CopeData$analysed_sst_15dy <- CopeData$analysed_sst_15dy-273.15
CopeData
```

Add Habitat column
```{r}
CopeData$Habitat <- ifelse(CopeData$depth >= -3000, "Shallow", "Deep")
```

Calanus pacificus linear model
```{r}
CalPacDF<- CopeData %>% select(c("SST", "Calanus.pacificus","date","Habitat","region"))

```

Looking at range of occurrence temps
```{r}
CalPacDFNoZero <- CalPacDF[CalPacDF$Calanus.pacificus != 0, ]

max(CalPacDFNoZero$SST)
#16.06244
min(CalPacDFNoZero$SST)
#3.517956
```


Transform count data
```{r}
CalPacDF$Calanus <- log10(CalPacDF$Calanus.pacificus+1)
str(CalPacDF)
CalPacDF
```

Temp only
```{r}
CalTemplm = lm(Calanus~SST, data=CalPacDF)
vif(CalTemplm)
summary(CalTemplm)
AIC(CalTemplm)
lrtest(CalTemplm)
```

Date only
```{r}
CalDatelm = lm(Calanus~date, data=CalPacDF)
vif(CalDatelm)
summary(CalDatelm)
AIC(CalDatelm)
lrtest(CalDatelm)
```

Depth
```{r}
CalDepthlm = lm(Calanus~Habitat, data=CalPacDF)
vif(CalDepthlm)
summary(CalDepthlm)
AIC(CalDepthlm)
lrtest(CalDepthlm)
```

Set Region 4 as the reference region
```{r}
levelCalPacDF <- within(CalPacDF, region <- relevel(region, ref = 4))
```

Region
```{r}
Calregionlm = lm(Calanus~region, data=levelCalPacDF)
vif(Calregionlm)
summary(Calregionlm)
AIC(Calregionlm)
lrtest(CalDepthlm,Calregionlm)
```

Temp and date
```{r}
CalTempDatelm = lm(Calanus~date+SST, data=CalPacDF)
vif(CalTempDatelm)
summary(CalTempDatelm)
AIC(CalTempDatelm)
lrtest(CalTemplm,CalTempDatelm)
lrtest(CalDatelm,CalTempDatelm)
#report this one in chart
```

Temp, date, and depth
```{r}
CalTempDateDepthlm = lm(Calanus~date+SST+Habitat, data=CalPacDF)
summary(CalTempDateDepthlm)
vif(CalTempDateDepthlm)
AIC(CalTempDateDepthlm)
lrtest(CalTempDatelm,CalTempDateDepthlm)
```

Temp, date, and region
```{r}
CalTempDateregionlm = lm(Calanus~SST+date+region, data=levelCalPacDF)
vif(CalTempDateregionlm)
summary(CalTempDateregionlm)
AIC(CalTempDateregionlm)
lrtest(CalTempDateDepthlm,CalTempDateregionlm)
```

Temp and Depth
```{r}
CalTempDepthlm = lm(Calanus~SST+Habitat, data=CalPacDF)
summary(CalTempDepthlm)
vif(CalTempDepthlm)
AIC(CalTempDepthlm)
lrtest(CalTemplm,CalTempDepthlm)
#22.855
lrtest(CalDepthlm,CalTempDepthlm)
#39.338
```

Set Region 4 as the reference region
```{r}
levelCalPacDF <- within(CalPacDF, region <- relevel(region, ref = 4))
```
Temp and Region
```{r}
CalTempRegionlm = lm(Calanus~SST+region, data=CalPacDF)
summary(CalTempRegionlm)
vif(CalTempRegionlm)
AIC(CalTempRegionlm)
lrtest(CalTempDepthlm,CalTempRegionlm)
plot(CalTempRegionlm)
```


Oithona spp. linear model 

```{r}
OithonaDF<- CopeData %>% select(c("SST", "Oithona.spp.","date","Habitat", "region"))

```

```{r}
OithinaDFNoZero <- OithonaDF[OithonaDF$Oithona.spp. != 0, ]

max(OithinaDFNoZero$SST)
#15.75873
min(OithinaDFNoZero$SST)
#3.304963
```

Transform count data
```{r}
OithonaDF$Oithona <- log10(OithonaDF$Oithona.spp.+1)
str(OithonaDF)
OithonaDF
```

Temp
```{r}
OiTemplm = lm(Oithona~SST, data=OithonaDF)
vif(OiTemplm)
summary(OiTemplm)
AIC(OiTemplm)
lrtest(OiTemplm)
```

Date only
```{r}
OiDatelm = lm(Oithona~date, data=OithonaDF)
summary(OiDatelm)
vif(OiDatelm)
AIC(OiDatelm)
lrtest(OiDatelm)
```

Depth
```{r}
OiDepthlm = lm(Oithona~Habitat, data=OithonaDF)
vif(OiDepthlm)
summary(OiDepthlm)
AIC(OiDepthlm)
lrtest(OiDepthlm)
```

Set Region 4 as the reference region
```{r}
OithonaLevelDF <- within(OithonaDF, region <- relevel(region, ref = 4))
```

Region
```{r}
Oiregionlm = lm(Oithona~region, data=OithonaLevelDF)
vif(Oiregionlm)
summary(Oiregionlm)
AIC(Oiregionlm)
lrtest(Oiregionlm)
```

Temp and date
```{r}
OiTempDatelm = lm(Oithona~date+SST, data=OithonaDF)
summary(OiTempDatelm)
vif(OiTempDatelm)
AIC(OiTempDatelm)
lrtest(OiTemplm,OiTempDatelm)
```


Temp, date, and depth
```{r}
OiTempDateDepthlm = lm(Oithona~date+SST+Habitat, data=OithonaDF)
summary(OiTempDateDepthlm)
vif(OiTempDateDepthlm)
AIC(OiTempDateDepthlm)
lrtest(OiTempDatelm,OiTempDateDepthlm)
```

Temp, date, and region
```{r}
OiTempDateregionlm = lm(Oithona~SST+date+region, data=OithonaLevelDF)
vif(OiTempDateregionlm)
summary(OiTempDateregionlm)
AIC(OiTempDateregionlm)
lrtest(OiTempDateDepthlm,OiTempDateregionlm)
```


Temp and region
```{r}
OiTempRegionlm = lm(Oithona~SST+region, data=OithonaLevelDF)
vif(OiTempRegionlm)
summary(OiTempRegionlm)
AIC(OiTempRegionlm)
lrtest(OiTemplm,OiTempRegionlm)
#12.588
lrtest(Oiregionlm,OiTempRegionlm)
#5.8846
lrtest(OiTempDateregionlm,OiTempRegionlm)
#2.2505
plot(OiTempRegionlm)

```


N. plumchrus linear model
```{r}
NeoDF<- CopeData %>% select(c("SST", "Neocalanus.plumchrus.V","date","Habitat","region"))

```

```{r}
NeoDFNoZero <- NeoDF[NeoDF$Neocalanus.plumchrus.V != 0, ]

max(NeoDFNoZero$SST)
#14.35687
min(NeoDFNoZero$SST)
#3.213618
```



Transform count data
```{r}
NeoDF$Neocalanus <- log10(NeoDF$Neocalanus.plumchrus.V+1)
str(NeoDF)
NeoDF
```

Temp
```{r}
NeoTemplm = lm(Neocalanus~SST, data=NeoDF)
vif(NeoTemplm)
summary(NeoTemplm)
AIC(NeoTemplm)
lrtest(NeoTemplm)
```

Date only
```{r}
NeoDatelm = lm(Neocalanus~date, data=NeoDF)
vif(NeoDatelm)
summary(NeoDatelm)
AIC(NeoDatelm)
lrtest(NeoDatelm)
```

Depth
```{r}
NeoDepthlm = lm(Neocalanus~Habitat, data=NeoDF)
vif(NeoDepthlm)
summary(NeoDepthlm)
AIC(NeoDepthlm)
lrtest(NeoDepthlm)
```

Set Region 4 as the reference region
```{r}
NeoLevelDF <- within(NeoDF, region <- relevel(region, ref = 4))
```

region
```{r}
Neoregionlm = lm(Neocalanus~region, data=NeoLevelDF)
vif(Neoregionlm)
summary(Neoregionlm)
AIC(Neoregionlm)
lrtest(NeoDepthlm,Neoregionlm)
```

Temp and date
```{r}
NeoTempDatelm = lm(Neocalanus~date+SST, data=NeoDF)
vif(NeoTempDatelm)
summary(NeoTempDatelm)
AIC(NeoTempDatelm)
lrtest(NeoTemplm,NeoTempDatelm)
lrtest(NeoDatelm,NeoTempDatelm)
```


Temp, date, and depth
```{r}
NeoTempDateDepthlm = lm(Neocalanus~date+SST+Habitat, data=NeoDF)
plot(NeoTempDateDepthlm)
vif(NeoTempDateDepthlm)
summary(NeoTempDateDepthlm)
AIC(NeoTempDateDepthlm)
lrtest(NeoTempDatelm,NeoTempDateDepthlm)
```

Temp, date, and region
```{r}
NeoTempDateregionlm = lm(Neocalanus~SST+date+region, data=NeoLevelDF)
vif(NeoTempDateregionlm)
summary(NeoTempDateregionlm)
AIC(NeoTempDateregionlm)
lrtest(NeoTempDateDepthlm,NeoTempDateregionlm)
```


Load in data and libraries for biomass peak analysis

HADLEY SST - Mean annual temp for each year
```{r}
BmpData <- read.csv(url("https://raw.githubusercontent.com/marisolgr/NEP_Copepods_ThermalPatterns/main/BiomassPeak.csv"))

head(BmpData)
tail(BmpData)

TempData <- read.csv(url("https://raw.githubusercontent.com/marisolgr/NEP_Copepods_ThermalPatterns/main/HadSST.csv"))
head(TempData)
tail(TempData)
library(dplyr)
library(ggplot2)
```

Annual avg
```{r}
TempData
library(dplyr)

tapply(TempData$SST, TempData$Year, mean)

AnnAvg <- c(9.43,9.1,9.86,10,10.1,10.3,9.5,9.12,8.68,9.18,9.43,9.07,8.91,9.84,10.8,11.1,10.7,9.81,10.4)

LagAnnAvg <- c(8.95,9.43,9.1,9.86,10,10.1,10.3,9.5,9.12,8.68,9.18,9.43,9.07,8.91,9.84,10.8,11.1,10.7,9.81)

AnnAvgData <- cbind(BmpData,AnnAvg)
head(AnnAvgData)
AnnAvgLagData <- cbind(BmpData,LagAnnAvg)
head(AnnAvgLagData)
```

```{r}
AnnWidthlm <- lm(CohortWidth~AnnAvg, data=AnnAvgData)
summary(AnnWidthlm)

AnnMidlm <- lm(Midpoint~AnnAvg, data=AnnAvgData)
summary(AnnMidlm)
```
No significant effect of temp on width (p=0.452), but significant effect of temp on midpoint (p=0.019)


Looking back at Batten and Mackas 2009 they used a one year lag - try this
```{r}
LagWidthlm <- lm(CohortWidth~LagAnnAvg, data=AnnAvgLagData)
summary(LagWidthlm)

LagMidlm <- lm(Midpoint~LagAnnAvg, data=AnnAvgLagData)
summary(LagMidlm)
```
no significance for lag ann avg and width p=0.20, but significance for midpoint p=0.006


Plot data

```{r}
lb1 <- paste("R^2 == ", "0.33")
lb2 <- paste("R^2 == ", "0.04")
```


```{r}
library(gridExtra)
library(ggpmisc)
plot1 <- ggplot(AnnAvgLagData, aes(x = LagAnnAvg))
  plot2 <- plot1 + geom_point(aes(y = Midpoint),size=3)+geom_smooth(method='lm', formula= y~x,  se=FALSE, color="black", aes(y = Midpoint))+   geom_text(x = 10, y = 155,
            label = lb1, parse = TRUE) + labs (y="Midpoint \n (day of year)", x="Annual Average SST - One Year Lag (°C)", tag="A")+ theme_bw(base_size=15)
  
  # adding the relative humidity data, transformed to match roughly the range of the temperature
  plot3 <- plot1 + geom_point(aes(y = CohortWidth),size=3)+geom_smooth(method='lm', formula= y~x,  se=FALSE, color="black", aes(y = CohortWidth))+  geom_text(x = 10, y = 50,
            label = lb2, parse = TRUE) + labs (y="Cohort Width \n (days)", x="Annual Average SST - One Year Lag (°C)", tag="B")+ theme_bw(base_size=15)
  
  grid.arrange(plot2, plot3, nrow = 2)
```


Looking at plumchrus abundance for blob years
```{r}
mhwNeoDF<- CopeData %>% select(c("SST", "Neocalanus.plumchrus.V","date","Habitat","region","year"))
tail(mhwNeoDF) 
mhwNeoDF$MHW <- ifelse(mhwNeoDF$year >= 2014, "MHW", "Amb")
write.csv(mhwNeoDF,"~/Desktop/gradschool/Internship/Practice", row.names = FALSE)
mhwNeoDF$Neocalanus <- log10(mhwNeoDF$Neocalanus.plumchrus.V+1)
aovmod <- aov(Neocalanus~MHW, mhwNeoDF)
summary(aovmod)
TukeyHSD(aovmod)
```

Let's try plotting our results and see what they look like
```{r}
PlotNeoDF<- mhwNeoDF %>% select(c("Neocalanus.plumchrus.V", "MHW"))
PlotNeoDF
ggplot(PlotNeoDF,aes(x=MHW,y=Neocalanus.plumchrus.V))+ 
  geom_boxplot(color="black")+xlab("Marine Heat Wave")+labs(y=expression(paste(italic("N. plumchrus")," Abundance")))
```

Try plotting as violin plot
```{r}
library(sinaplot)
PlotNeoDF
ggplot(PlotNeoDF, aes(MHW, Neocalanus.plumchrus.V)) +
  geom_violin(aes(fill=MHW))+scale_fill_manual(values=c("dodgerblue4","red4") ) +xlab("Marine Heat Wave")+labs(y=expression(paste(italic("N. plumchrus")," Abundance")))
```


